---
title: "data exploration"
author: "XIAO MA"
date: "12/12/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(ggplot2) 
#install.packages("ggExtra")
library(ggExtra)
#install.packages("GGally")
library(GGally) 
library(dplyr)
library(skimr)
```

## Import the data set 
```{r}
cdi = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() 

# Create a new variable called crm_1000, which is the crime rate per 1000 population in each county in year 1990, and another variable poparea which is the population density per square mile. Also change the number of doctors and beds into doctors and beds per 1000 population.

cdi = cdi %>% 
  mutate(
    poparea = pop/area,
    crm_1000 = (crimes/pop)*1000,
    docs = (docs/pop) * 1000,
    beds = (beds/pop) * 1000
  )
```

## To preview the data
```{r}
glimpse(cdi)
summary(cdi)

```

## Draw pairwise plot to check association betwee every two variable
```{r}
ggpairs(cdi[,-c(1,2,3,10)])
```

From the pairwise plot, we observed the correlation and distribution between every two variables. The association between crime rate to each variable was displayed in the last row. No strong linear relationship was detected between the crime rate per 1000 person and other variables. Between variables, we noticed that Percent of population aged 18-34 have negative relation with Percent of population aged 65+. Although the high absolute value of correlation indicates a strong relationship, it may not be linear relation.

## Marginal 
```{r}
p = ggplot(cdi, aes(x=pop, y=crimes)) +
      geom_point() +
      theme(legend.position="none")
p1 = ggMarginal(p, type = "density")
p1

p2 = ggplot(cdi, aes(x=totalinc, y = pop)) +
      geom_point() +
      theme(legend.position="none")
p2 = ggMarginal(p2, type = "density")
p2

#not ob linear
p3 = ggplot(cdi, aes(x = pop18, y = pop65)) + geom_point() +
      theme(legend.position="none")
p3 = ggMarginal(p3, type = "density")
p3

#not ob linear
p4 = ggplot(cdi, aes(x = hsgrad, y = bagrad)) + geom_point() +
      theme(legend.position="none")
p4 = ggMarginal(p4, type = "density")
p4


p5 = ggplot(cdi, aes(x = hsgrad, y = poverty)) + geom_point() +
      theme(legend.position="none")
p5 = ggMarginal(p5, type = "density")
p5



p6 = ggplot(cdi, aes(x = hsgrad, y = unemp)) + geom_point() +
      theme(legend.position="none")
p6 = ggMarginal(p6, type = "density")
p6

p7 = ggplot(cdi, aes(x = hsgrad, y = pcincome)) + geom_point() +
      theme(legend.position="none")
p7 = ggMarginal(p7, type = "density")
p7

p8 = ggplot(cdi, aes(x = bagrad, y = pcincome)) + geom_point() +
      theme(legend.position="none")
p8 = ggMarginal(p8, type = "density")
p8

p9 = ggplot(cdi, aes(x = bagrad, y = unemp)) + geom_point() +
      theme(legend.position="none")
p9 = ggMarginal(p9, type = "density")
p9

p10 = ggplot(cdi, aes(x = bagrad, y = poverty)) + geom_point() +
      theme(legend.position="none")
p10 = ggMarginal(p10, type = "density")
p10

```



## Crime Rate Comparison Across Region
```{r}
#CRM_1000 vs. Geographic Region
cdi_region = cdi %>%
  group_by(region) %>%
  summarise(region_crm_1000 = sum(crimes)*1000/sum(pop)) %>%
  arrange(desc(region_crm_1000))

cdi_region = cdi_region %>%
  mutate(region_cla = case_when(
  region == 1 ~ "Northeast",
  region == 2 ~ "North Central",
  region == 3 ~ "South",
  region == 4 ~ "West")) %>% 
  mutate(region_cla = as.factor(region_cla)) %>% 
  arrange(region_cla,region_crm_1000)

ggplot(cdi_region) + 
  geom_bar(stat = "identity",aes( y = region_crm_1000,reorder(region_cla,+region_crm_1000)),fill = "darkblue",width = 0.3) +
  labs(x = "Region", y = "Crime Rate",title = "The Crime Rate of Each Geographic Region") +
  theme(plot.title = element_text(hjust = 0.5))
```

Southern US have highest crime rate. 

## State crime( find the state with unusual crime rate)
```{r}
#CRM_1000 vs. state
cdi_state = cdi %>%
  group_by(state) %>%
  summarise(state_crm_1000 = sum(crimes)*1000/sum(pop)) %>%
  arrange(desc(state_crm_1000))

theme_dotplot <- theme_bw(10) +
    theme(axis.text.y = element_text(size = rel(0.75)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(0.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5),
        panel.grid.minor.x = element_blank())

ggplot(cdi_state,aes(state_crm_1000,reorder(state,state_crm_1000))) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(30, 110),breaks = seq(30,110,5)) + 
  theme_dotplot +
  labs(y = "State",title = 
         "The Crime Rate with different States") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(cdi_state,aes(y = state_crm_1000)) + geom_boxplot()
```

According to the boxplot, there is no outliers, which means there is no states with unusual crime rate.
```{r}
cdi_desc = cdi %>%
  arrange(desc(crm_1000))
head(cdi_desc)
ggplot(cdi_desc,aes(y = crm_1000)) + geom_boxplot()
```

There are four county with unusual crime rate, and they are (Kings, NY), (123 St._Loui, MO),(70, Fulton, GA), (9 Dade, FL)


